---
title: "City vs Cancellation CODE"
author: "Group 50"
date: "12/1/2023"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
flights_2018_2023 <- read.csv("C:/Users/yaroa/Downloads/flights_2018_2023_top_airports.csv")  #TO BE ADJUSTED TO WHERE DATA IS STORED ON PERSONAL DEVICE FOR NEW USER

data = flights_2018_2023

#install.packages("vctrs")

library(foreign)
library(dplyr)
library(missForest)
library(leaps)
library(xtable)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
#library(caret)
library(ggplot2)
library(rattle)
library(maptree)

```

```{r}
#Cleaning the Data
summary(data)
```

```{r}
unique_cities = unique(data$ORIGIN_CITY_NAME)
unique_cities
```



```{r}
#subset$education = as.factor(subset$education)
#data = subset(data, select = -c(BRANDED_CODE_SHARE) ) #No need to know if code shared
#data = subset(data, select = -c(MKT_CARRIER_AIRLINE_ID, MKT_UNIQUE_CARRIER)) #No need to know other MKT info (all same)
#data = subset(data, select = -c(SCH_OP_UNIQUE_CARRIER, SCH_OP_CARRIER_AIRLINE_ID, SCH_OP_CARRIER)) #Not needed

data_cancel = subset(data, select = c(YEAR, MONTH, DAY_OF_MONTH, DAY_OF_WEEK, MKT_CARRIER, ORIGIN_AIRPORT_ID, ORIGIN, ORIGIN_CITY_NAME, ORIGIN_STATE_ABR, DEST_AIRPORT_ID, DEST, DEST_CITY_NAME, ARR_DELAY_NEW, DEST_STATE_ABR, ARR_DEL15, DIVERTED, CARRIER_DELAY, WEATHER_DELAY, NAS_DELAY, SECURITY_DELAY, LATE_AIRCRAFT_DELAY, CANCELLED, ORIGIN_CITY_MARKET_ID))
                     
                     
data_cancel$YEAR = as.integer(data_cancel$YEAR)
data_cancel$MONTH = as.factor(data_cancel$MONTH)
data_cancel$DAY_OF_MONTH = as.integer(data_cancel$DAY_OF_MONTH)
data_cancel$DAY_OF_WEEK = as.factor(data_cancel$DAY_OF_WEEK)
data_cancel$MKT_CARRIER = as.factor(data_cancel$MKT_CARRIER)
data_cancel$ORIGIN_AIRPORT_ID = as.factor(data_cancel$ORIGIN_AIRPORT_ID)
data_cancel$ORIGIN = as.factor(data_cancel$ORIGIN)
data_cancel$ORIGIN_CITY_NAME = as.factor(data_cancel$ORIGIN_CITY_NAME)
data_cancel$ORIGIN_STATE_ABR = as.factor(data_cancel$ORIGIN_STATE_ABR)
data_cancel$DEST_AIRPORT_ID = as.factor(data_cancel$DEST_AIRPORT_ID)
data_cancel$DEST = as.factor(data_cancel$DEST)
data_cancel$DEST_CITY_NAME = as.factor(data_cancel$DEST_CITY_NAME)
data_cancel$DEST_STATE_ABR = as.factor(data_cancel$DEST_STATE_ABR)
data_cancel$ARR_DEL15 = as.factor(data_cancel$ARR_DEL15)
data_cancel$DIVERTED = as.factor(data_cancel$DIVERTED)
data_cancel$CARRIER_DELAY = as.factor(data_cancel$CARRIER_DELAY)
data_cancel$WEATHER_DELAY = as.factor(data_cancel$WEATHER_DELAY)
data_cancel$NAS_DELAY = as.factor(data_cancel$NAS_DELAY)
data_cancel$SECURITY_DELAY = as.factor(data_cancel$SECURITY_DELAY)
data_cancel$LATE_AIRCRAFT_DELAY = as.factor(data_cancel$LATE_AIRCRAFT_DELAY)
data_cancel$CANCELLED = as.factor(data_cancel$CANCELLED)

summary(data_cancel)

```

# CANCELLATION FOR CHICAGO

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Chicago, IL")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .001))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```




# CANCELLATION FOR NEW YORK

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "New York, NY")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .001))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```



# CANCELLATION FOR Newark, NJ

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Newark, NJ")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 

summary(data_cancel_city$CANCELLED)
9148 / nrow(data_cancel_city)
```




# CANCELLATION FOR San Francisco, CA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "San Francisco, CA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```





# CANCELLATION FOR Denver, CO

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Denver, CO")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```






# CANCELLATION FOR Long Beach, CA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Long Beach, CA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```




# CANCELLATION FOR Newburgh/Pughkeepsie, NY

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Newburgh/Poughkeepsie, NY")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .001))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```




# CANCELLATION FOR Los Angeles, CA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Los Angeles, CA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```






# CANCELLATION FOR Dallas/Fort WOrth, TX

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Dallas/Fort Worth, TX")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .001))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```




# CANCELLATION FOR Santa Ana, CA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Santa Ana, CA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .0015))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```





# CANCELLATION FOR Phoenix, AZ

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Phoenix, AZ")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .0018, minbucket = 90))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```







# CANCELLATION FOR Fort Lauderdale, FL

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Fort Lauderdale, FL")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```




# CANCELLATION FOR White Plains, NY

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "White Plains, NY")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```


# CANCELLATION FOR Islip, NY

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Islip, NY")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .0015))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```



# CANCELLATION FOR Charlotte, NC

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Charlotte, NC")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```









# CANCELLATION FOR Las Vegas, NV

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Las Vegas, NV")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```




# CANCELLATION FOR Houston, TX

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Houston, TX")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```





# CANCELLATION FOR Orlando, FL

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Orlando, FL")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```




# CANCELLATION FOR Atlanta, GA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Atlanta, GA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```





# CANCELLATION FOR Burbank, CA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Burbank, CA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```



# CANCELLATION FOR Miami, FL

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Miami, FL")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```



# CANCELLATION FOR San Jose, CA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "San Jose, CA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```





# CANCELLATION FOR Ontario, CA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Ontario, CA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```


# CANCELLATION FOR Oakland, CA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Oakland, CA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```



# CANCELLATION FOR Dallas, TX

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Dallas, TX")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```


# CANCELLATION FOR Seattle, WA

```{r}
#Splitting data into different departure cities
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_NAME == "Seattle, WA")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .00097))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin City =", as.character(high)), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```










# TREES FOR MARKETS
```{r}
markets = unique(data_cancel$ORIGIN_CITY_MARKET_ID)
markets
```

# CANCELLATION FOR New York City Area

```{r}
#Splitting data into different departure cities
#31703 32575 31057 32467 30559 30194 32211 32457 31453 30977 30466 31454 30325 30397
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_MARKET_ID == "31703")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + ORIGIN + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002, minbucket = 100))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin Market = New York City Area"), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```



# CANCELLATION FOR Southern California

```{r}
#Splitting data into different departure cities
#31703 32575 31057 32467 30559 30194 32211 32457 31453 30977 30466 31454 30325 30397
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_MARKET_ID == "32575")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + ORIGIN + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002, minbucket = 100))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin Market = Southern California Area"), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```





# CANCELLATION FOR Charlotte, NC Area

```{r}
#Splitting data into different departure cities
#31703 32575 31057 32467 30559 30194 32211 32457 31453 30977 30466 31454 30325 30397
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_MARKET_ID == "31057")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + ORIGIN + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002, minbucket = 100))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin Market = Charlotte, NC Area"), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```







# CANCELLATION FOR  Southern Florida Area

```{r}
#Splitting data into different departure cities
#31703 32575 31057 32467 30559 30194 32211 32457 31453 30977 30466 31454 30325 30397
data_cancel_city = data_cancel[c(which(data_cancel$ORIGIN_CITY_MARKET_ID == "32467")),]
#summary(data_cancel_city)

#TREE
set.seed(716)
tree = rpart(CANCELLED ~ DEST + ORIGIN + MONTH + YEAR + DAY_OF_WEEK + MKT_CARRIER + DAY_OF_WEEK + ORIGIN, data = data_cancel_city, method = "class", control = rpart.control(cp = .002, minbucket = 100))
high = data_cancel_city$ORIGIN_CITY_NAME[1]
fancyRpartPlot(tree, main = paste("Single Tree for Origin Market = Southern Florida Area"), palettes=c("Greens", "Reds"))


#Obtain the fitted probabilities
probs3 <- predict(tree)
probs3 <- as.matrix(probs3)

#Obtain the Predictions
preds3 = apply(probs3, 1, function(x) which.max(x))
preds3 = ifelse(preds3 ==1, 0,1)

#Creates a confusion matrix
as.matrix(table(preds3, data_cancel_city$CANCELLED) ) ##this means that we misclassified

#Train CER
1 - (sum(preds3 == data_cancel_city$CANCELLED)/nrow(data_cancel_city)) 
```




THERE IS NOT A SIGNIFICANT DIFFERENCE BETWEEN AIRPORTS WITHIN 1 MARKET VALUE WHEN FITTING THE TREES

















